\documentclass[11pt,fullpage]{article}
\usepackage[urlcolor=blue,colorlinks=true]{hyperref}

\oddsidemargin 0.0in
\textwidth 6.5in

\begin{document}

\title{Genome Analysis Toolkit (GATK) User Manual}
\author{Matt Hanna}
\date{30 Mar 2009}
\maketitle

\section{Walkers}
\subsection{Command-line Arguments}
Users can create command-line arguments for walkers by creating public
member variables annotated with @Argument in the walker.  The
@Argument annotation takes a number of different parameters:

\begin{enumerate}
  \item [fullName] The full name of this argument.  Defaults to the
    toLowerCase()'d member name.  When specifying fullName on the 
    command line, prefix a double dash (--).
  \item [shortName] The alternate, short name for this argument.
    Defaults to the first letter of the member name.  When specifying 
    shortName on the command line, prefix a single dash (-).
  \item [doc] Documentation for this argument.  Will appear in help
    output when a user either requests help with the --help (-h)
    argument or when a user specifies an invalid set of arguments.
  \item [required] Whether the argument is required when used with
    this walker.  Default is required = true.
  \item [exclusive] Specifies that this argument is mutually
    exclusive of another argument in the same walker.  Defaults to not
    mutually exclusive of any other arguments.
  \item [defaultValue] Specifies the default value for this parameter,
    in string form.  
\end{enumerate}

Examples:

Create an required int parameter with full name --myint, short name
-m.
Pass this argument by adding ``--myint 6'' or -m6 to the command line.
\begin{verbatim}
@Argument
public int myInt;
\end{verbatim}

Create an optional float parameter with full name
--myFloatingPointArgument, short name -m.  Pass this argument by
adding --myFloatingPointArgument 2.71 or -m2.71.
\begin{verbatim}
@Argument(fullName="myFloatingPointArgument",required=false,defaultValue="3.14159")
public float myFloat;
\end{verbatim}

The GATK will parse the argument differently depending on the type of
the public member variable's type.  Many different argument types are 
supported, including primitives and their wrappers, arrays, typed and 
untyped collections, and any type with a String constructor.

When the GATK cannot completely infer the type (such as in the case of
untyped collections), it will assume that the argument is a String.
GATK is aware of concrete implementations of some interfaces and
abstract classes.  If the argument's member variable is of type List
or Set, the GATK will fill the member variable with a concrete
ArrayList or TreeSet, respectively.  Maps are not currently supported.

\subsection{Output}
By default, the walkers provide protected out and err PrintStreams.
Users can write to these streams just as they write to System.out and
System.err.  This output can be redirected to a file using the out,
err, and outerr command-line arguments.

\end{document}
